---
title: "simpleRegress"
author: "Giselle Gonzalez"
date: "Summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

 for model length~neck, adjusted r^2 = 0.75
 
## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## length based on age

```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```


adjuste $r^2$ for the age model is 0.51, this is less than the 
adjuste r^2 for neck, so this model is not as good as the neck model



## length based on chest

```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```

 The model length based on chest is better than adjusteD $r^2$ for the age model (0.51) and adjuste r^2 for neck model(0.75), because r^2 for chest is (0.79)

## length based on weight

```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)
```

this model length based weight is not as good as the model -length based on chest because the adjusted r-squared for length based on weight is lower(0.74) than the adjusted r-squared for chest(0.79)


## length based on sex 

```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)
```
  this model is the lowest model  because the rsquare for length based on sex is (-0.0) this means its lower than chest r-squared(0.79)
  
  
## length based on headlength

```{r}
headlenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenModel)
summary.lm(headlenModel)
```

This model length based on headlength is way better than length based on chest because the r-squared for chest is (0.78)
 and which is lower than the r-squared for headlen which was (0.84)

## length based on headwth

```{r}
headwthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwthModel)
summary.lm(headwthModel)
```

This model is not better than the model headlen because the r^2 for headwth is lower(0.53) and the headlen was (0.84) which is higher 

## conclusion

I will be putting all the models from least to the greatest value of the adjusted r-squared
~least 
length based on sex is (-0.0)  
length based on age (0.51)  
LENgth based on headwth (0.53)  
length based weight(0.74)  
length~neck (0.75)  
length based on chest (0.79)
LENgth based on headlength (0.84)
~greatest