---
title: "simpleRegress"
author: "Giselle Gonzalez"
date: "Summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

 for model length~neck, adjusted r^2 = 0.75
 
## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## length based on age

```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```


adjuste $r^2$ for the age model is 0.51, this is less than the 
adjuste r^2 for neck, so this model is not as good as the neck model



## length based on chest

```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```

 The model length based on chest is better than adjuste $r^2$ for the age model (0.51) and adjuste r^2 for neck model(0.75), because r^2 for chest is (0.79)

## length based on weight

```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)
```

this model length based weight is not as good as the model -length based on chest because the adjusted r-squared for length based on weight is lower(0.74) than the adjusted r-squared for chest(0.79)


## length based on sex 

```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)
```
  this model is not better than the chest model because the rsquare for length based on sex is (-0.0) this means its lower than chest r-squared(0.79)
